---
title: "Learn_ggplot2"
author: "GINNY"
date: "25 May 2017"
output: html_document
---


```{r}

  setwd("/Users/ginny/Downloads/Rgraphics")
  
  housing = read.csv("dataSets/landdata-states.csv")
  
  head(housing [1:5])
  
  
#####################################simple plots  
  hist(housing$Home.Value)
  
  ###using ggplot2
  
  
  library(ggplot2)
  
  ggplot(housing, aes(x=Home.Value))+
    geom_histogram()
  
  
######################################complex plots
  
  
  plot(Home.Value ~ Date,
       data=subset(housing, State == "MA"))
  points(Home.Value ~ Date, col="red",
         data=subset(housing, State == "TX"))
  legend(1975, 400000,
         c("MA", "TX"), title="State",
         col=c("black", "red"),
         pch=c(1, 1))
  
  ###using ggplot2
  
  
  ggplot(subset(housing, State %in% c("MA","TX")),
         aes(x=Date, y=Home.Value, color=State))+
    geom_point()
  
```

#Getting to know the basics
```{R}
## aesthetic mapping
## geometic objects
  
help.search("geom_",package="ggplot2")  
  
```

## Points

```{R}
hp2001Q1=subset(housing, Date==2001.25)
  
ggplot(hp2001Q1,
       aes(y=Structure.Cost, x=Land.Value))+
  geom_point()
  
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = log(Land.Value))) +
  geom_point()

```
## Lines
```{R}
hp2001Q1$pred.SC =predict(lm(Structure.Cost ~ log(Land.Value), data=hp2001Q1))

p1= ggplot(hp2001Q1, aes(x=log(Land.Value), y=Structure.Cost))

p1+geom_point(aes(color=Home.Value))+geom_line(aes(y=pred.SC))

####the attributes in aes() should be subset of the dataframe's attributes

```


## Smoothers

```{R}
p1+geom_point(aes(color=Home.Value))+
  geom_smooth()


```


##Text

```{R}
p1+geom_text(aes(label=State),size=3)

library(ggrepel)

p1+
  geom_point()+
  geom_text_repel(aes(label=State),size=3)


```

## Aesthetic mapping and assignment

Variables are mapped to aesthetics with the aes() function, while fixed aesthetics are set outside the aes() call.

```{R}

p1+
  geom_point(aes(size=2),
    color="red")
```



##Mapping Variables To other Aesthetics


```{R}
p1+
  geom_point(aes(color=Home.Value, shape=region))
```


## Exercise 1

```{R}

  setwd("/Users/ginny/Downloads/Rgraphics")

dat=read.csv("dataSets/EconomistData.csv")


```

### Create a scatter plot with CPI on the x axis and HDI on the y axis
```{R}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()

```

### Color the points blue
```{R}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(color="blue")

```

### Map the color of the points to Region

```{R}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(aes(color=Region))

```

### Make the points bigger by setting size to 2

```{R}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(aes(color=Region),size=2)

```


### Map the size of the points to HDI.Rank

```{R}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(aes(color=Region,size=HDI.Rank))

```

#Statistical Transformations

boxplots, histograms and prediction lines etc require statistical transformation

for boxplot, the y value must be transformed to the median and 1.5

```{r}
args(geom_histogram)
args(stat_bin)

```


##Setting statistical transformation arguments

```{R}

p2=ggplot(housing, aes(x=Home.Value))

p2+geom_histogram()

p2+geom_histogram(stat="bin", binwidth = 4000)


```

###changing the statistical transformation 
```{r}
housing.sum=aggregate(housing["Home.Value"],by=housing["State"], FUN=mean)
rbind(head(housing.sum),tail(housing.sum))


ggplot(housing.sum, aes(x=State, y=Home.Value))+
  geom_bar(stat="identity")



```

##Exercise II

### Re-create a scatter plot with CPI on the x axis and HDI on the y axis (as you did in the previous exercise)

```{R}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()

```


### Overaly a smoothing line on top of the scatter plot using geom_smooth

```{R}
p3=ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()

p3+
  geom_smooth()

```


### Overlay a smoothing line on top of the scatter plot using geom_smooth, but use a linear model for the predictions. Hint: see ?stat_smooth


```{R}
p3=ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()

p3+
  geom_smooth(method="lm")

```

### Overlay a smoothing line on top of the scatter plot using geom_line, hint: change the statistical transformation.


```{R}
p3=ggplot(dat, aes(x = CPI, y = HDI)) 


dat$ps<- predict(lm(HDI ~ CPI, data = dat))


p3 + 
  geom_point() +
  geom_line(aes(y=dat$ps))

```

### Overlay a smoothing line on top of the scatter plot using the default loess method, but make it less smooth. Hint: see ?loess

```{R}
p3=ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()

p3+
  geom_smooth(span=0.2)+
  geom_smooth(span=1,color="red")

```

#scales
to control aesthetic mapping

name, limits, breaks, labels

##dotplot example

```{R}
p4=ggplot(housing,
          aes(x=State,
              y=Home.Price.Index))+
  theme(legend.position="top",
        axis.text=element_text(size=6))

p5=p4+geom_point(aes(color=Date),
                 alpha=0.5,  ###alpha is the transparency value
                 size=1.5,
                 position=position_jitter(width=0.1,height =0))



p5+scale_x_discrete(name="State Abbreviation")+
  scale_color_continuous(name="",
                         breaks=c(1976,1994,2013),
                         labels=c("'76","'94","'13"))


p5+scale_x_discrete(name="State Abbreviation")+
  scale_color_continuous(name="",
                         breaks=c(1976,1994,2013),
                         labels=c("'76","'94","'13"),
                         low="blue",high="red")


library(reshape2)
library(scales)

p5+scale_x_discrete(name="State Abbreviation")+
  scale_color_continuous(name="",
                         breaks=c(1976,1994,2013),
                         labels=c("'76","'94","'13"),
                         low=muted("blue"),high=muted("red"))


```

##Exercise III


### create a scatter plot with CPI on the x axis and HDI on the y axis, color the points to indicate region 

```{R}
p6=ggplot(dat, aes(x=CPI,y=HDI))


p6+
  geom_point(aes(color=Region))
```

### Modify the x,y and color scales so that they have more easily-understood names

```{R}
p6=ggplot(dat, aes(x=CPI,y=HDI))


p6+
  geom_point(aes(color=Region))+
  scale_x_discrete(name="Corruption Perception Index")+
  scale_y_discrete(name="Human Development Index")

```


###Modify the color scale to use specific values of your choosing 
```{R}
p6=ggplot(dat, aes(x=CPI,y=HDI))


p6+
  geom_point(aes(color=Region))+
  scale_x_discrete(name="Corruption Perception Index")+
  scale_y_discrete(name="Human Development Index")+
  scale_color_manual(values=c("red","blue","green","black","grey","orange"))

```

# Faceting

faceting is ggplot2 parlance for small multiples

the idea is to create seperate graphs for subsets of data

`facet_wrap` and `facet_grid`

facilites comparision among plots, not just geoms within a plot



```{r}
p_5=ggplot(housing, aes(x=Date, y=Home.Value))

p_5+geom_line(aes(color=State))

```

two problems here, too many states to distinguish by color and the lines obscure one another


```{r}
p_5=ggplot(housing, aes(x=Date, y=Home.Value))

p_5+geom_line()+facet_wrap(~State,ncol=10)

```


# Themes

```{r}
p_5=ggplot(housing, aes(x=Date, y=Home.Value))+geom_line()+facet_wrap(~State,ncol=10)
p_5+theme_linedraw()

p_5+theme_light()

```


## Overiding theme defaults

```{r}
p_5=ggplot(housing, aes(x=Date, y=Home.Value))+geom_line()+facet_wrap(~State,ncol=10)

p_5+theme_minimal()+
  theme(text=element_text(color="turquoise"))


```

#wrong and right

```{r}
housing.byyear=aggregate(cbind(Home.Value, Land.Value)~ Date, data=housing, mean)

ggplot(housing.byyear,
       aes(x=Date))+
  geom_line(aes(y=Home.Value), color="red")+
  geom_line(aes(y=Land.Value), color="blue")

###right version

library(tidyr)

home.land.byyear=gather(housing.byyear, value="value", key="type",
                        Home.Value, Land.Value)

###add one more variable called type.

ggplot(home.land.byyear, aes(x=Date, y=value,
                             color=type))+
  geom_line()



####a bit of tidyr

library(dplyr)

stocks=data_frame(time=as.Date('2009-01-01')+0:9,
                  X=rnorm(10,0,1),
                  Y=rnorm(10,0,2),
                  Z=rnorm(10,0,4))


gather(stocks, stock, price, -time)


tid=stocks %>%gather(stock, price, -time)  ###-time meaning get everything other than time


#putting it all together




```

```{r}
###it seems r square should be calculated by myself

library(ggrepel)

model=lm(HDI~poly(CPI,2),  dat)

dat$pred=predict(model, x=dat$CPI)
sm=summary(model)
r_s=sm$r.squared
dat$label=dat$Country

important_countries=c("Italy","France","Greece","Russia","Spain", "Norway","New Zealand","US","China","South Africa","Venezuela","Congo","Rwanda","India","Botswana","Germany","Singapore","Japan","Sudan","Iraq","Brazil","Afghamstan","Myanmar","Bhutan","Britain")

muted_which=which(dat$Country%in%important_countries==F)

dat$label[muted_which]=""

p6=ggplot(dat, aes(x=CPI,y=HDI))

p6+
  geom_line(aes(y=pred),size=1,color="red")+
  geom_point(aes(color=Region),size=4,shape=21,stroke=1.5,fill="white")+
  geom_text_repel(aes(label=label), size = 4,point.padding = unit(0.3,"lines"))+
  scale_color_manual(values=c("dodgerblue","darkturquoise","seagreen","deepskyblue4","brown1","brown"))+
  scale_x_continuous(name="Corruption Perception Index",limits = c(1,10),breaks=seq(1:10))+
  scale_y_continuous(name="Human Development Index",limits=c(0.2,1.0))+
  theme(legend.position = "top" ,
        legend.key = element_rect(fill="white"),
        legend.title = element_blank(),
        legend.text = element_text(size=12),
        axis.title = element_text(size=12),
        axis.text = element_text(size=10),
        panel.background = element_rect(colour = "white",fill="white"),
        panel.grid.major.y =element_line(size = 0.1,colour = "grey"),
        plot.title = element_text(face="bold",size=16, hjust = 0))+
        guides(colour = guide_legend(nrow = 1))+
  ggtitle("Corruption and human development")

  



```



